---
title: "Install archived Bioconductor packages from previous releases"
author: "Marcel Ramos"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  BiocStyle::html_document:
    number_sections: yes
    toc: true
vignette: >
  %\VignetteIndexEntry{BiocArchive Quick Start}
  %\VignetteEnginve{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

`BiocArchive` is a package dedicated to preserving reproducibility with older
Bioconductor versions. It only works on older Bioconductor releases such
as `3.14`. Note that users must have the proper `3.14` setup to be able
to install packages from the archive. This means that users should be running
R version `4.1`.

It is highly recommended that users run docker containers with the appropriate
R version installation and install the package via GitHub or via source.

# Installation

Currently, it is available via GitHub. 

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("LiNk-NY/BiocArchive")
```

# Last release version

The `lastBuilt` helper function finds the last built date for the supplied
Bioconductor version. Other functions rely on this date to install the
appropriate packages.

```{r}
lastBuilt(version = "3.14")
```

# Archived installations

To install Bioconductor packages, it is minimally required that the user
install via `BiocManager` on the docker container. The added value that
BiocArchive provides is that it provides versioned installations of CRAN
packages from either the RStudio Package Manager (`RSPM`) or the Microsoft R
Archive Network (`MRAN`).

Bioconductor installations should work as normal as if one were installing the
package via `BiocManager` within the legacy `3.14` container.

```{r}
install("DESeq2", version = "3.14", dry.run = TRUE)

install("MultiAssayExperiment", version = "3.14", dry.run = TRUE)
```

# CRAN installations from the source archive

Packages on CRAN have a history of versions at a particular URL location:

<https://cran.r-project.org/src/contrib/Archive>

A CRAN package from the archive can be installed using `CRANinstall`:

```{r}
CRANinstall("dplyr", "3.14", dry.run = TRUE)
```

Note that this method only works on one package at a time. Future support
will allow installations for multiple packages at a time. 

# Session Information

```{r}
sessionInfo()
```


